---
title: "Project2"
author: "LiChia Chang"
date: "2023-10-07"
output: 
  github_document:
    toc: true
    # html_preview: True
---
# Introduction
This document is a vignette to show how to retrieve data from an API. To showcase this, I will create multiple functions for refining API endpoints and design helper functions to enhance the user-friendliness of the API. In the final section, I will employ these functions to extract data that I'm interested in and conduct exploratory data analysis.

This article dives into the [AirVisual API](https://www.iqair.com/commercial-air-quality-monitors/api), which collects data from the largest network of ground-based sensors worldwide, providing the most accurate and reliable air quality information. Further details will be discussed in the subsequent sections.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Requirements
To interact with the AirVisual API, you will requrie several libraries. In the R environment console, you can install these libraries by using the following command: `install.packages(package_name)`. Make sure to replace package_name with the name of any library from the list below that you haven't installed in your environment.


```{r message=FALSE}
library(httr)
library(dplyr)
library(jsonlite)
library(ggplot2)
```


# API Interaction Functions
In this section, I will demonstrate how to establish contact with the AirVisual API and develop functions for specific purposes. I will provide separate explanations for each function, detailing their usage and intended purpose.

*Note: For this project, I am utilizing my own API key, which allows for up to 10,000 calls per month. If you wish to use your own API key, simply replace "key=" with your key. You can find detailed information regarding API key usage at this website: [https://www.iqair.com/dashboard/api](https://www.iqair.com/dashboard/api)*


* __Get all supported countries__

```{r}
get_country <- function(){
  url <- 'http://api.airvisual.com/v2/countries?key=b94cdf68-3b7b-4808-9012-ffb6359d7690'
  response <- GET(url)
  country_list <- content(response, encoding = "UTF-8")$data
  
  ## Convert to data frame
  results <- data.frame(country = sapply(country_list, '[[', 'country'))
  return(results)
}
```


* Verify if the country exists in the supported country list

Verify if the country is in the supported list

```{r}
verify_country <- function(country="USA"){
  
  url <- 'http://api.airvisual.com/v2/countries?key=b94cdf68-3b7b-4808-9012-ffb6359d7690'
  
  response <- GET(url)
  country_list <- content(response, encoding = "UTF-8")$data
  
  for (c in country_list) {
     if (country == c) {
       return(TRUE)
     }
  }
  return(FALSE)
}
```


* Get all supported states inside a specified country

```{r}
get_state <- function(country="USA"){
  
  custom_country <- gsub(" ", "%20", country)
  
  url <- "http://api.airvisual.com/v2/states?country="
  str_key <- "&key=b94cdf68-3b7b-4808-9012-ffb6359d7690"
  url <- paste0(url, custom_country, str_key)
  
  
  response <- GET(url)
  state_list <- content(response, encoding = "UTF-8")$data
  
  ## Convert to data frame
  results <- data.frame(country = sapply(state_list, '[[', 'state'))
  return(results)
}
```


* Verify supported states in a country

Verify if the state is in the supported list of specific country

```{r}
verify_state <- function(country="USA", state="California"){
  
  custom_country <- gsub(" ", "%20", country)
  
  url <- "http://api.airvisual.com/v2/states?country="
  str_key <- "&key=b94cdf68-3b7b-4808-9012-ffb6359d7690"
  url <- paste0(url, custom_country, str_key)
  
  
  response <- GET(url)
  state_list <- content(response, encoding = "UTF-8")$data
  
  for (s in state_list) {
     if (state == s) {
       return(TRUE)
     }
  }
  return(FALSE)
}
```


* Get all supported cities inside a specified country and state

```{r}
get_city <- function(country="USA", state="California"){

  custom_state <- gsub(" ", "%20", state)
  custom_country <- gsub(" ", "%20", country)
  
  url <- "http://api.airvisual.com/v2/cities?state="
  str_country <- "&country="
  str_key <- "&key=b94cdf68-3b7b-4808-9012-ffb6359d7690"
  url <- paste0(url, custom_state, str_country, custom_country, str_key)
  
  
  response <- GET(url)
  city_list <- content(response, encoding = "UTF-8")$data
  
  ## Convert to data frame
  results <- data.frame(country = sapply(city_list, '[[', 'city'))
  return(results)
}
```


* Verify supported cities in a state

Verify if the city is in the supported list of specific country and state

```{r}
verify_city <- function(country="USA", state="California", city="Los Angeles"){

  custom_state <- gsub(" ", "%20", state)
  custom_country <- gsub(" ", "%20", country)
  
  url <- "http://api.airvisual.com/v2/cities?state="
  str_country <- "&country="
  str_key <- "&key=b94cdf68-3b7b-4808-9012-ffb6359d7690"
  url <- paste0(url, custom_state, str_country, custom_country, str_key)
  
  
  response <- GET(url)
  city_list <- content(response, encoding = "UTF-8")$data
  
  for (cty in city_list) {
     if (city == cty) {
       return(TRUE)
     }
  }
  return(FALSE)
}
```


* Verify if the connection is sucessfully connected
return status of connection
```{r}
test_connection <- function(country="USA", state="California", city="Los Angeles") {
  ## replace " " to "%20", which is a URL-encoded representation of a space character.
  custom_city <- gsub(" ", "%20", city)
  custom_state <- gsub(" ", "%20", state)
  custom_country <- gsub(" ", "%20", country)
  
  url <- 'http://api.airvisual.com/v2/city?city='
  str_state <- '&state='
  str_country <- '&country='
  str_key <- '&key=b94cdf68-3b7b-4808-9012-ffb6359d7690'
  url <- paste0(url, custom_city, str_state, custom_state, str_country, custom_country, str_key)
  
  
  response <- GET(url)
  status <- content(response)$status
  return(status)
}
```


* Helper function to wrap up all the validation

```{r}
wrap_validation <- function(country="USA", state="California", city="Los Angeles") {
  
  status <- test_connection(country, state, city)
  if (status != "success") {
    msg <- "ERROR: Connection failed"
    stop(msg)
  }
  
  if (!verify_country(country)) {
    msg <- "Not a valid country in supported list"
    stop(msg)
  }
  
  if (!verify_state(country, state)) {
    msg <- paste0("Not a valid state in the supported list of country: ", country)
    stop(msg)
  }
  
  if (!verify_city(country, state, city)) {
    msg <- paste0("Not a valid city in the supported list of country: ", country, " and state: ", state)
    stop(msg)
  }
  
  print("Everything works well")
}
```


* Get data object
Return the most recent data object by given state, city, and country

```{r}
get_data <- function(city="Los Angeles", state="California", country="USA"){

  ## replace " " to "%20", which is a URL-encoded representation of a space character.
  custom_city <- gsub(" ", "%20", city)
  custom_state <- gsub(" ", "%20", state)
  custom_country <- gsub(" ", "%20", country)
  
  
  url <- 'http://api.airvisual.com/v2/city?city='
  str_state <- '&state='
  str_country <- '&country='
  str_key <- '&key=b94cdf68-3b7b-4808-9012-ffb6359d7690'
  url <- paste0(url, custom_city, str_state, custom_state, str_country, custom_country, str_key)
  
  
  response <- GET(url)
  results <- content(response, encoding = "UTF-8")$data
  return(results)
}
```


* Get current weather
Return the most recent weather string by given state, city, and country

```{r}
get_weather <- function(city="Los Angeles", state="California", country="USA"){

  ## replace " " to "%20", which is a URL-encoded representation of a space character.
  custom_city <- gsub(" ", "%20", city)
  custom_state <- gsub(" ", "%20", state)
  custom_country <- gsub(" ", "%20", country)
  
  
  url <- 'http://api.airvisual.com/v2/city?city='
  str_state <- '&state='
  str_country <- '&country='
  str_key <- '&key=b94cdf68-3b7b-4808-9012-ffb6359d7690'
  url <- paste0(url, custom_city, str_state, custom_state, str_country, custom_country, str_key)
  
  
  response <- GET(url)
  ic <- content(response, encoding = "UTF-8")$data$current$weather$ic
  
  if (ic == "01d") {
    weather <- "clear sky (day)"
  }
  if (ic == "01n") {
    weather <- "clear sky (night)"
  }
  if (ic == "02d") {
    weather <- "few clouds (day)"
  }
  if (ic == "02n") {
    weather <- "few clouds (night)"
  }
  if (ic == "03d") {
    weather <- "scattered clouds"
  }
  if (ic == "04d") {
    weather <- "broken clouds"
  }
  if (ic == "09d") {
    weather <- "shower rain"
  }
  if (ic == "10d") {
    weather <- "rain (day time)"
  }
  if (ic == "10n") {
    weather <- "rain (night time)"
  }
  if (ic == "11d") {
    weather <- "thunderstorm"
  }
  if (ic == "13d") {
    weather <- "snow"
  }
  if (ic == "50d") {
    weather <- "mist"
  }
  
  return(weather)
}
```


# Exploratory Data Analysis

study the weather and temperature in north carolina state
```{r}
city_list <- get_city(country="USA", state="North Carolina")
city_list
```

We want to see the relationship between coordinates and pollution
```{r}
show_city <- c("Raleigh", "Cary", "Apex", "Durham", "Chapel Hill", "Garner")

show_city_list_NC <- data.frame(country = show_city)

```

```{r results = FALSE}
df <- data.frame()
row_index <- 1

apply(show_city_list_NC, MARGIN = 1, FUN = function(x) {
  Sys.sleep(20)
  
  city_data <- get_data(country="USA", state="North Carolina", city = x)
  latitude <- city_data$location$coordinates[[1]]
  longitude <- city_data$location$coordinates[[2]]
  pollution <- city_data$current$pollution$aqius
  temperature <- city_data$current$weather$tp
  humidity <- city_data$current$weather$hu
  
  tmp <- data.frame(city = x, lat = latitude, lon = longitude, pol = pollution, temp = temperature, hu = humidity)
  
   # Set the row index as the row names of tmp
  rownames(tmp) <- row_index
  
  # Add tmp as a new row to the global df data frame
  df <<- rbind(df, tmp)  # Use '<<-' to assign to the global variable
  
  # Increment the row index for the next row
  row_index <<- row_index + 1
  
})
```


```{r}
ggplot(df, aes(x = lat, y = lon, color = pol)) +
  geom_point() +
  geom_text(aes(label = city), vjust = 1, hjust = 1) + 
  labs(x = "latitude", y = "longitude", color = "pollution (AQI)") + 
  scale_color_gradient(low = "green", high = "blue")
```

```{r}
ggplot(df, aes(x = lat, y = lon, color = temp)) +
  geom_point() +
  geom_text(aes(label = city), vjust = 1, hjust = 1) + 
  labs(x = "latitude", y = "longitude", color = "temperature (Celsius)") + 
  scale_color_gradient(low = "green", high = "blue")
```

```{r}
ggplot(df, aes(x = lat, y = lon, color = hu)) +
  geom_point() +
  geom_text(aes(label = city), vjust = 1, hjust = 1) + 
  labs(x = "latitude", y = "longitude", color = "pollution (AQI)") + 
  scale_color_gradient(low = "green", high = "blue")
```

